Fail Faster

### 2025-03-23 paper/paper.tex
typederived

\subsubsection{\scalaname}
In \scalaname, we implement a subset of the \camlname
strategies: the boolean list generator, single-pass BST generator, and
well-typed STLC term generator\footnote{By default, ScalaCheck does not support deriving gnenerators from types,
so those strategies are excluded.}. We did not implement an optimized version of ScalaCheck's
\rand, but our experimental setup was otherwise the same as in \secref{subsubsection:ocaml}.

@@

We run each
strategy on all tasks using a 60-second timeout.  If a bug is not found within
this limit, the task is considered unsolved.  Tasks where all strategies fail to
find the bug are excluded from our dataset, as they provide no basis for
comparison.  Similarly, we exclude tasks where the BQ strategy completes in
under 5ms, as such negligible runtimes do not yield meaningful insights into
relative performance. Across all 30 seeds, these filters remove
28/600 (4.67\%) of type-derived STLC's tasks,
168/600 (28\%) of STLC's tasks,
149/1110 (14.42\%) of type-derived BST's tasks,
268/1110 (24.1\%) of single-pass BST's tasks, and
371/1110 (33.42\%) of insertion-based BST's tasks. More sophisticated
strategies tend to find bugs faster, leading to a higher number of filtered
tasks. By applying these filters, we ensure that our reported speedups reflect
optimizations that meaningfully impact performance.

Our results are summarized in \figref{fig:etna}, which shows the geometric average of 
individual-task speedups for each strategy and benchmark. Trends in bug-finding speed reflect trends in performance
from \secref{subsection:benchmarking}. The distribution of speedups in \figref{fig:swarm} reveals
substantial variability, with most tasks clustering near the median and a long tail of outliers
achieving much larger gains. STLC shows a more bimodal distribution of speedups
than the other benchmarks, which we attribute to the heterogenous difficulty of its tasks:
some tasks regularly hit the 60-second timeout, while others finish in a fraction of a second. We
find that ``easier'' tasks---those that run on the order of milliseconds---regularly achieve speedups in the
range of $4-9\times$, whereas tasks that run on the order of seconds achieve more moderate speedups of up to $3.5\times$.
Overall,
these results give strong evidence that \camlname consistently benefits bug-finding speed, sometimes drastically.
\begin{figure}[h]
  \includegraphics[scale=0.3]{speedup.png}
  \caption{Geometric average of all speedups---relative to BQ---for each strategy and benchmark, showing that staging leads to better bug-finding speed across the board.
  }
  \label{fig:etna}
\end{figure}
\begin{figure}[h]
  \includegraphics[scale=0.31]{swarm.png}
  \caption{Distribution of individual-task speedups across strategies and benchmarks. Gray dots represent tasks. Plots with extremely low variance have no swarm overlay; in plots with high variance, outliers correspond to large speedups.}
  \label{fig:swarm}
\end{figure}

\section{Beyond Strict Functional Languages}
\label{section:other-langs}

>>>>

\subsubsection{\scalaname}
In \scalaname, we implement a subset of the \camlname
strategies: the boolean list generator, single-pass BST generator, and
well-typed STLC term generator.\footnote{By default, ScalaCheck does not support deriving gnenerators from types,
so those strategies are excluded.} We did not implement an optimized version of ScalaCheck's
\rand, but our experimental setup was otherwise the same as in \secref{subsubsection:ocaml}.

@@

We run each
strategy on all tasks using a 60-second timeout.  If a bug is not found within
this limit, the task is considered unsolved.  Tasks where all strategies fail to
find the bug are excluded from our dataset, as they provide no basis for
comparison.  Similarly, we exclude tasks where the \bq strategy completes in
under 5ms, as such negligible runtimes do not yield meaningful insights into
relative performance. Across all 30 seeds, these filters remove
28/600 (4.67\%) of type-derived STLC's tasks,
168/600 (28\%) of STLC's tasks,
149/1110 (14.42\%) of type-derived BST's tasks,
268/1110 (24.1\%) of single-pass BST's tasks, and
371/1110 (33.42\%) of insertion-based BST's tasks. More sophisticated
strategies tend to find bugs faster, leading to a higher number of filtered
tasks. By applying these filters, we ensure that our reported speedups reflect
optimizations that meaningfully impact performance.

\begin{figure}[h]
  \includegraphics[scale=0.3]{speedup.png}
  \caption{Geometric average of all speedups---relative to \bq---for each strategy and benchmark, showing that staging leads to better bug-finding speed across the board.
  }
  \label{fig:etna}
\end{figure}

Our results are summarized in \figref{fig:etna}, which shows the geometric average of 
individual-task speedups for each strategy and benchmark. Trends in bug-finding speed reflect trends in performance
from \secref{subsection:benchmarking}. The distribution of speedups in \figref{fig:swarm} reveals
substantial variability, with most tasks clustering near the median and a long tail of outliers
achieving much larger gains. STLC shows a more bimodal distribution of speedups
than the other benchmarks, which we attribute to the heterogenous difficulty of its tasks:
some tasks regularly hit the 60-second timeout, while others finish in a fraction of a second. We
find that ``easier'' tasks---those that run on the order of milliseconds---regularly achieve speedups in the
range of $4-9\times$, whereas tasks that run on the order of seconds achieve more moderate speedups of up to $3.5\times$.
Overall,
these results give strong evidence that \camlname consistently benefits bug-finding speed, sometimes drastically.

\begin{figure}[h]
  \includegraphics[scale=0.31]{swarm.png}
  \caption{Distribution of individual-task speedups across strategies and benchmarks. Gray dots represent tasks. Plots with extremely low variance have no swarm overlay.}
  \label{fig:swarm}
\end{figure}

\subsubsection{Do staged type-derived generators reduce timeouts?}
In addition to \camlname's overall bug-finding ability, we evaluate 
the performance of its type-derived generators---specifically, 
whether they can transform previously intractable tasks into 
tractable ones. Out of 78 tasks that initially timed out, 
50 (64\%) completed successfully after our interventions. 
This suggests that staged type-derived generators can provide 
significant performance improvements for free---that is, without 
requiring any additional effort from the developer.
\section{Beyond Strict Functional Languages}
\label{section:other-langs}
